\chapter*{Summary}
In recent years, all areas of society have been affected by an unprecedented process of data digitization and a huge increment in computational capacity. Cross-sectional areas of knowledge are the ones that have benefited the most from this so-called \emph{Fourth Industrial Revolution}. This is specially true for robotics, as most of its applications must be embedded in devices with limited resources and must work in real-time.

One of the most relevant fields of study within this area is \emph{Human-Robot Interaction}, as it is a core component for systems in which the robot needs to be aware of the humans around for planning its actions. Owing to the nature of our own forms of communication, audio and video signals are usually employed to perform such task, but the latter has the advantage of giving the robot relevant information without explicit interaction.

Our work focuses on human pose estimation, a classic computer vision problem closely related with \emph{Human-Robot Interaction}. This task has been historically addressed by means of cumbersome motion capture systems, but huge efforts have been made in the last decades in order to develop algorithms that can solve it making use of \emph{off-the-shelf} devices. The inclusion of deep learning techniques has rocketed the effectiveness of such methods in just a few years. However, the amount of labeled data and computational resources required might be prohibitive, specially when dealing with three-dimensional data, which provides the robot with more extensive information about its surroundings.

Taking all of these challenges into account, we propose an \emph{end-to-end} pipeline for estimating 3D human poses that works in real-time in an \emph{off-the-shelf} computer with a commercial RGBD sensor. In that regard, our main contributions are the performance of a quantitative evaluation of 2D pose estimators in the literature, the design and development of the proposed system, a quantitative evaluation and comparison of this system against a \emph{state-of-the art} solution for 3D pose estimation and a real-time demo that experimentally validates its performance in qualitative terms.

The obtained results reveal that our solution is very competitive in terms of accuracy, while being lightweight enough to run in real-time, as we demonstrate in both our quantitative and qualitative evaluations. Therefore, our system is suitable for being integrated in robotics applications in which the robot is required to interact with humans.

